\documentstyle[11pt]{article}
% Default margins are too wide.
\setlength{\topmargin}{-.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{.125in}
\setlength{\textwidth}{6.25in}
\begin{document}
\title{CS 6316 -- Midterm}
\author{Donnie Newell}
\date{March 6, 2013}
\maketitle

{\bf Part 1 - Short Answer}
\begin{enumerate}
	\item
	Stock Predictor
	\begin{enumerate}
		\item
		ease of understanding

		\item
		speed of predictions

		\item
		easily update models with new data

	\end{enumerate}
	\item
	choose from multiple consistent hypotheses

	\item
	growth of hypothesis space

	\item
	Machine Learning Bias

	\item
	noisy data
	\begin{enumerate}
		\item
		Candidate elimination
		
		\item
		Decision Tree
		
		\item
		Explanation
		
	\end{enumerate}
	
	\item
	predict disease with attributes of varying cost
	\begin{enumerate}
		\item
		which classification algorithm?
		
		\item
		why?
		
	\end{enumerate}
	
	\item
	perfect classification
	
	\item
	ID3 decision tree
	
	\item
	2nd mystery data set
	\begin{enumerate}
		\item
		most likely class using naive bayes w/o smoothing?
		
		\item
		naive bayes with laplace smoothing(k = 1) on prior and class-conditional parameters
		
	\end{enumerate}
		
	\item
	Bill goes to the hospital
	\begin{enumerate}
		\item
		Draw Bayesian network
		
		\item
		joint probability distribution as the product of conditional probabilities
		
		\item
		independent parameters needed
		
		\item
		w/o conditional independence between variables, how many independent parameters are needed?
		
		\item
		Markov blanket of $T_2$
	\end{enumerate}
	\item
	Bayesian Networks
	\begin{enumerate}
		\item
		
		\item
		
	\end{enumerate}
		
\end{enumerate}

{\bf Part 2 - Experiments}
\begin{enumerate}
	\item
	Compare Naive Bayes and kNN
	\begin{enumerate}
		\item
		naive bayes vs. iB1 using training set
		\begin{enumerate}
			\item
			which is better?
			
			\item
			why?

		\end{enumerate}
		
		\item
		repeat previous using 10-fold cross-val.
		\begin{enumerate}
			\item
			what changed?
			
			\item
			why?
			
		\end{enumerate}
	\end{enumerate}
	
	\item
	Analyze kNN
	\begin{enumerate}
		\item
		Fill out table for 4 models
		
		\item
		Analysis
		\begin{enumerate}
			\item
			effect of distance weighting function on relative classification accuracy
			
			\item
			difference between numeric and nominal attributes
			
			\item
			results with respect to $k$
		\end{enumerate}
		
	\end{enumerate}
	
\end{enumerate}
\end{document}